Understanding the Ratings of Children's Books: A Data-Driven Analysis

### 2024-12-04 paper/paper.qmd
improve the qmd file and references file, render the pdf

---
title: "Understanding the Ratings of Children's Books: A Data-Driven Analysis"
subtitle: "Exploring the Impact of Cover Type, Page Count, and Publication Period on Children's Book Ratings"
author: Zien Gao
thanks: "Code and data supporting this analysis are available at: <https://github.com/lauragao75/Understanding-the-Popularity-of-Children-s-Books-A-Data-Driven-Analysis>"
date: today
date-format: long
abstract: "This study investigates the factors influencing children's book ratings, focusing on attributes such as cover type, page count, and publication period. Using simulated and cleaned datasets from various sources, we applied a Bayesian hierarchical linear model to explore these relationships comprehensively. We estimate that hardcover books, books in the 150-299 and 300-499 page ranges, and books published after 2000 have notably higher ratings, providing a clear indication of preferences in children's literature. These insights are critical for publishers, authors, and parents in making informed decisions about book design and selection. The findings highlight important characteristics that contribute to well-received children's literature, aiming to enhance both quality and reader satisfaction."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

#### Workspace setup ####
library(broom)
library(tidyverse)
library(dplyr)
library(knitr)
library(ggplot2)
library(here)
library(kableExtra)
library(arrow)
library(modelsummary)

@@

## Overview

@@

The libraries used in our analysis include `tidyverse` [@tidyverse], `arrow` [@arrow], `broom` [@broom], `ggplot2` [@ggplot2], `knitr` [@knitr], `kableExtra` [@kableExtra], `here` [@here], and `modelsummary` [@modelsummary]. These libraries enable data manipulation, model building, and visualization, ensuring a comprehensive examination of the dataset to explore relationships between book characteristics and ratings.

@@

```{r}
#| label: fig-ratings-distribution
#| fig-cap: Distribution of Children's Book Ratings. Shows how the ratings of children's books are generally skewed towards higher values, indicating a general satisfaction with the literature available.
#| echo: false

@@

@fig-ratings-distribution shows the distribution of ratings across all books in the dataset. We observe that most books tend to have relatively high ratings, clustering around values between 3.5 and 5. This skew towards higher ratings may indicate a general satisfaction with the books or potential bias in how ratings are assigned.

@@

```{r}
#| label: fig-cover-ratings
#| fig-cap: Relationship Between Book Ratings and Cover Type. Illustrates that hardcover books tend to receive slightly higher ratings compared to paperback editions, potentially due to their perceived durability and premium quality.
#| echo: false

@@

```{r}
#| label: fig-page-range-ratings
#| fig-cap: Average Ratings by Page Range. Indicates that books with 150-299 pages are more likely to receive higher ratings, potentially reflecting a balance between story depth and accessibility for children.
#| echo: false

@@

```{r}
#| label: fig-pub-period-ratings
#| fig-cap: Average Ratings by Publication Period. Highlights that books published after 2000 tend to have higher ratings, possibly reflecting evolving preferences and improved production standards.
#| echo: false

@@

Overall, we observe interesting patterns between book attributes and ratings. The physical characteristics of books, such as cover type and length, appear to play a role in how well they are rated. Similarly, the time of publication also seems to impact the perceived quality of a book. These insights suggest that certain characteristics may enhance a book's appeal, providing valuable information for authors and publishers seeking to create well-received children’s books.

@@

Define \(y_i\) as the average rating for book \(i\). Then \(eta_1\) represents the effect of cover type, \(eta_2\) represents the effect of page range, and \(eta_3\) represents the effect of publication period.

```{=tex}
\[
egin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + eta_1(CoverType_i) + eta_2(PageRange_i) + eta_3(PublicationPeriod_i) \\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
eta_1 &\sim \mbox{Normal}(0, 2.5) \\
eta_2 &\sim \mbox{Normal}(0, 2.5) \\
eta_3 &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}
\]
```

@@

To provide context for the model results, we present some key summary statistics of the dataset used in the analysis. This includes the average rating, distribution of book types, and the breakdown of page ranges and publication periods.

```{r}
#| label: tbl-summary-statistics
#| echo: false
#| tbl-cap: "Summary statistics of the children's book dataset"

library(dplyr)
summary_stats <- analysis_data %>%
  summarise(
    `Average Rating` = round(mean(rating, na.rm = TRUE), 2),
    `Number of Hardcover Books` = sum(cover == "Hardcover"),
    `Number of Paperback Books` = sum(cover == "Paperback")
  )

# Get counts for page ranges and publication periods
pages_range_count <- analysis_data %>%
  count(pages_range) %>%
  rename(Metric = pages_range, Value = n) %>%
  mutate(Metric = paste("Page Range:", Metric))

publish_period_count <- analysis_data %>%
  count(publish_period) %>%
  rename(Metric = publish_period, Value = n) %>%
  mutate(Metric = paste("Publication Period:", Metric))

# Combine all summary statistics into one table
combined_summary <- bind_rows(
  summary_stats %>% pivot_longer(everything(), names_to = "Metric", values_to = "Value"),
  pages_range_count,
  publish_period_count
)

# Display the combined summary statistics table
knitr::kable(combined_summary, caption = "Combined summary statistics of the children's book dataset, including average ratings, distribution of cover types, page ranges, and publication periods.")
```

@@

\newpage
\appendix

@@

```{r}
#| label: tbl-regression-results
#| echo: false
#| tbl-cap: "Regression model summary for predicting children's book ratings, including estimated coefficients and credible intervals."

@@

To ensure that the Bayesian hierarchical linear model has converged properly, we used multiple convergence diagnostics. The primary convergence diagnostic used is the **Gelman-Rubin statistic** (R-hat). The R-hat value should ideally be close to 1, indicating that the Markov chains have mixed well and reached convergence. Any R-hat value significantly above 1 suggests non-convergence and needs further investigation.

```{r}
#| label: rhat-diagnostics
#| echo: false
#| fig-cap: "Gelman-Rubin convergence diagnostic (R-hat) values for each model parameter. All values are close to 1, indicating that the Markov chains have mixed well and achieved convergence."

@@

Another convergence diagnostic involves examining trace plots. Trace plots show the sampled values of each parameter across iterations for each Markov chain. Well-mixed chains with no clear trends suggest that the model has converged.

```{r}
#| label: trace-plot
#| echo: false
#| fig-cap: "Trace plots for key parameters across the Markov chains, showing well-mixed chains with stable values. The lack of discernible patterns or trends confirms that convergence has been reached."

# Trace plot for key parameters in the model
plot(children_model, plotfun = "trace")
```

@@

Posterior predictive checks are used to assess the goodness-of-fit of the Bayesian model. By simulating data from the posterior distribution, we can compare the simulated data to the observed data to determine if the model adequately represents the data.

```{r}
#| label: posterior-checks
#| echo: false
#| fig-cap: "Posterior predictive checks showing a comparison between the observed data and simulated data generated from the model's posterior distribution. This helps evaluate whether the model adequately captures the key features of the observed data."

@@

Below is a summary table of the key parameters in the model, including the estimated means, standard deviations, and 95% credible intervals.

```{r}
#| label: posterior-summary
#| echo: false
#| fig-cap: "Summary statistics of the posterior distributions for key model parameters, including posterior means, standard deviations, and 95% credible intervals, which provide insights into the central tendencies and uncertainties of the parameter estimates."

# Summary of the posterior distributions
summary(children_model)
```

@@

Posterior density plots provide a visual representation of the uncertainty in the model parameters. The density plots show the distribution of possible values for each parameter, which helps in understanding the spread and central tendency.

```{r}
#| label: fig-results-effects
#| echo: false
#| fig-cap: "Posterior density plots for each model parameter, providing a visual representation of the uncertainty in the parameter estimates. The density curves illustrate the range and central tendency of plausible values derived from the data and priors."

@@

## Data Collection and Challenges

@@

## Addressing Non-Response Bias

@@

- **Contact**: For any inquiries, please contact Zien Gao via lauragao75@gmail.com.

>>>>

---
title: "Understanding the Ratings of Children's Books: A Data-Driven Analysis"
subtitle: "Exploring the Impact of Cover Type, Page Count, and Publication Period on Children's Book Ratings"
author: Zien Gao
thanks: "Code and data supporting this analysis are available at: <https://github.com/lauragao75/Understanding-the-Ratings-of-Children-s-Books-A-Data-Driven-Analysis>"
date: today
date-format: long
abstract: "This study investigates the factors influencing children's book ratings, focusing on attributes such as cover type, page count, and publication period. Using simulated and cleaned datasets from various sources, we applied a Bayesian hierarchical linear model to explore these relationships comprehensively. We estimate that hardcover books, books in the 150-299 and 300-499 page ranges, and books published after 2000 have notably higher ratings, providing a clear indication of preferences in children's literature. These insights are critical for publishers, authors, and parents in making informed decisions about book design and selection. The findings highlight important characteristics that contribute to well-received children's literature, aiming to enhance both quality and reader satisfaction."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

#### Workspace setup ####
library(bayesplot)
library(broom)
library(tidyverse)
library(dplyr)
library(knitr)
library(ggplot2)
library(here)
library(kableExtra)
library(arrow)
library(modelsummary)

@@

@@

The libraries used in our analysis include `tidyverse` [@tidyverse], `arrow` [@arrow], `broom` [@broom], `ggplot2` [@ggplot2], `knitr` [@knitr], `kableExtra` [@kableExtra], `here` [@here], `bayesplot` [@bayesplot], `broom.mixed` [@broommixed], `coda` [@coda] and `modelsummary` [@modelsummary]. These libraries enable data manipulation, model building, and visualization, ensuring a comprehensive examination of the dataset to explore relationships between book characteristics and ratings.

@@

```{r}
#| label: fig-ratings-distribution
#| fig-cap: "Distribution of Children's Book Ratings. Shows how the ratings of children's books are generally skewed towards higher values, indicating a general satisfaction with the literature available."
#| echo: false
#| warning: false
#| message: false

@@

@fig-ratings-distribution shows the distribution of ratings across all books in the dataset. We observe that most books tend to have relatively high ratings, clustering around values between 3.5 and 4.5. This skew towards higher ratings may indicate a general satisfaction with the books or potential bias in how ratings are assigned.

@@

```{r}
#| label: fig-cover-ratings
#| fig-cap: "Relationship Between Book Ratings and Cover Type. Illustrates that hardcover books tend to receive slightly higher ratings compared to paperback editions, potentially due to their perceived durability and premium quality."
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-page-range-ratings
#| fig-cap: "Average Ratings by Page Range. Indicates that books with 150-299 pages are more likely to receive higher ratings, potentially reflecting a balance between story depth and accessibility for children."
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-pub-period-ratings
#| fig-cap: "Average Ratings by Publication Period. Highlights that books published after 2000 tend to have higher ratings, possibly reflecting evolving preferences and improved production standards."
#| echo: false
#| warning: false
#| message: false

@@

Overall, we observe interesting patterns between book attributes and ratings. The physical characteristics of books, such as cover type and length, appear to play a role in how well they are rated. Similarly, the time of publication also seems to impact the perceived quality of a book.

@@

Define $y_i$ as the average rating for book $i$. Then $\eta_1$ represents the effect of cover type, $\eta_2$ represents the effect of page range, and $\eta_3$ represents the effect of publication period.

\begin{align}
y_i \mid \mu_i, \sigma &\sim \text{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \eta_1(\text{CoverType}_i) + \eta_2(\text{PageRange}_i) + \eta_3(\text{PublicationPeriod}_i) \\
\alpha &\sim \text{Normal}(0, 2.5) \\
\eta_1 &\sim \text{Normal}(0, 2.5) \\
\eta_2 &\sim \text{Normal}(0, 2.5) \\
\eta_3 &\sim \text{Normal}(0, 2.5) \\
\sigma &\sim \text{Exponential}(1)
\end{align}

@@

To provide context for the model results, we present some key summary statistics of the dataset used in the analysis. @tbl-summary-table includes the average rating, distribution of book types, and the breakdown of page ranges and publication periods.

```{r}
#| label: tbl-summary-table
#| echo: false
#| tbl-cap: "Summary statistics of the children's book dataset, showing the distribution of key metrics such as average rating, book cover types, page ranges, and publication periods."

library(dplyr)
library(ggplot2)
library(knitr)
library(tidyr)

# Function to create count summaries
create_count_summary <- function(data, count_column, prefix) {
  data %>%
    count({{ count_column }}) %>%
    rename(Metric = {{ count_column }}, Value = n) %>%
    mutate(Metric = paste(prefix, Metric, sep = ": "))
}

# Summary statistics
summary_stats <- analysis_data %>%
  summarise(
    `Average Rating` = round(mean(rating, na.rm = TRUE), 2),
    `Number of Hardcover Books` = sum(cover == "Hardcover"),
    `Number of Paperback Books` = sum(cover == "Paperback")
  ) %>%
  pivot_longer(everything(), names_to = "Metric", values_to = "Value")

# Page range and publication period counts
pages_range_count <- create_count_summary(analysis_data, pages_range, "Page Range")
publish_period_count <- create_count_summary(analysis_data, publish_period, "Publication Period")

# Combine summary statistics
combined_summary <- bind_rows(summary_stats, pages_range_count, publish_period_count)

# Generate table
knitr::kable(
  combined_summary,
) %>%
  kable_styling(latex_options = "hold_position")

```

@@

\appendix

@@

@tbl-regression-results summarizes the regression results, showing the estimated coefficients, median absolute deviations, and credible intervals for each predictor.
```{r}
#| label: tbl-regression-results
#| echo: false
#| tbl-cap: "Regression model summary for predicting children's book ratings, including estimated coefficients and credible intervals."
#| warning: false
#| message: false

@@

To ensure that the Bayesian hierarchical linear model has converged properly, we used multiple convergence diagnostics. The primary convergence diagnostic used is the **Gelman-Rubin statistic** (R-hat). The R-hat value should ideally be close to 1, indicating that the Markov chains have mixed well and reached convergence. Any R-hat value significantly above 1 suggests non-convergence and needs further investigation. In @fig-rhat-diagnostics , we present the R-hat values for each model parameter.

```{r}
#| label: fig-rhat-diagnostics
#| echo: false
#| fig-cap: "Gelman-Rubin convergence diagnostic (R-hat) values for each model parameter. All values are close to 1, indicating that the Markov chains have mixed well and achieved convergence."

@@

Another convergence diagnostic involves examining trace plots. Trace plots show the sampled values of each parameter across iterations for each Markov chain. Well-mixed chains with no clear trends suggest that the model has converged. In @fig-trace-plot , we observe the trace plots for key parameters in the model.

```{r}
#| label: fig-trace-plot
#| echo: false
#| fig-cap: "Trace plots for key parameters across the Markov chains, showing well-mixed chains with stable values. The lack of discernible patterns or trends confirms that convergence has been reached."
#| fig-width: 10 
#| fig-height: 7 
#| dpi: 300

# Trace plot for key parameters in the model
plot(children_model, plotfun = "trace",)
```

@@

Posterior predictive checks are used to assess the goodness-of-fit of the Bayesian model. By simulating data from the posterior distribution, we can compare the simulated data to the observed data to determine if the model adequately represents the data. @fig-posterior-checks illustrates the results of posterior predictive checks for the Bayesian model.

```{r}
#| label: fig-posterior-checks
#| echo: false
#| fig-cap: "Posterior predictive checks showing a comparison between the observed data and simulated data generated from the model's posterior distribution. This helps evaluate whether the model adequately captures the key features of the observed data."

@@

@tbl-posterior-sum is a summary table of the key parameters in the model, including the estimated means, standard deviations, and 95% credible intervals.

```{r}
#| label: tbl-posterior-sum
#| echo: false
#| tbl-cap: "Summary statistics of the posterior distributions for key model parameters, including posterior means, standard deviations, and 95% credible intervals, which provide insights into the central tendencies and uncertainties of the parameter estimates."
#| warning: false
#| message: false

library(broom.mixed)
library(knitr)
library(kableExtra)
library(coda)

# Summarize the posterior distributions into a tidy data frame
posterior_summary <- broom.mixed::tidyMCMC(
  as.matrix(children_model), 
  conf.int = TRUE, 
  conf.level = 0.95
)

# Render the summary as a table
knitr::kable(
  posterior_summary
) %>%
  kable_styling(latex_options = "hold_position", full_width = FALSE)

```

@@

Posterior density plots provide a visual representation of the uncertainty in the model parameters. @fig-results-effects show the distribution of possible values for each parameter, which helps in understanding the spread and central tendency.

```{r}
#| label: fig-results-effects
#| echo: false
#| fig-cap: "Posterior density plots for each model parameter, providing a visual representation of the uncertainty in the parameter estimates. The density curves illustrate the range and central tendency of plausible values derived from the data and priors."
#| warning: false
#| fig-width: 10 
#| fig-height: 7 
#| dpi: 300

@@

## Data Collection Challenges and Bias

@@

@@
